Assessment of zonal symmetric and asymmetric components of the Southern Annular Mode using a novel approach

### 2020-10-10 analysis/paper/paper.Rmd
render

---
twocol: FALSE  #  Use FALSE for submission. 
title: Assessment of zonal symmetric and asymmetric components of the Southern Annular Mode using a novel approach
author1: 
  name: Elio Campitelli
  email: elio.campitelli@cima.fcen.uba.ar
author2: 
  name: Leandro B. Díaz
affiliation: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos, Buenos Aires, Argentina
    CONICET – Universidad de Buenos Aires, Centro de Investigaciones del Mar y la Atmósfera (CIMA), Buenos Aires, Argentina
    CNRS – IRD – CONICET – UBA, Instituto Franco‐Argentino para el Estudio del Clima y sus Impactos (UMI 3351 IFAECI), Buenos Aires, Argentina
extra_authors: 
  - name: Carolina Vera
    affiliation: |
     Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos, Buenos Aires, Argentina
     CONICET – Universidad de Buenos Aires, Centro de Investigaciones del Mar y la Atmósfera (CIMA), Buenos Aires, Argentina
     CNRS – IRD – CONICET – UBA, Instituto Franco‐Argentino para el Estudio del Clima y sus Impactos (UMI 3351 IFAECI), Buenos Aires, Argentina
abstract: |
  Enter the text of your abstract here. This is a sample American Meteorological Society (AMS) \LaTeX\ template.  This document provides authors with instructions on the use of the AMS \LaTeX\ template.  Authors should refer to the file amspaper.tex to review the actual \LaTeX\ code used to create this document. The template.tex file should be modified by authors for their own manuscript.
statement: |
  This is significant becasue I wrote it.
bibliography: ["AsymSAM.bib","packages.bib"]
output: 
  bookdown::pdf_book: 
    base_format: rticles::ams_article
    latex_engine: "xelatex"
  bookdown::word_document2: default
header-includes: 
 - \usepackage{gensymb}
 - \usepackage{subfig}
 - \usepackage[inline]{showlabels}
---

>>>>

---
twocol: FALSE  #  Use FALSE for submission. 
title: Assessment of zonal symmetric and asymmetric components of the Southern Annular Mode using a novel approach
author1: 
  name: Elio Campitelli
  email: elio.campitelli@cima.fcen.uba.ar
author2: 
  name: Leandro B. Díaz
affiliation: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos, Buenos Aires, Argentina
    CONICET – Universidad de Buenos Aires, Centro de Investigaciones del Mar y la Atmósfera (CIMA), Buenos Aires, Argentina
    CNRS – IRD – CONICET – UBA, Instituto Franco‐Argentino para el Estudio del Clima y sus Impactos (UMI 3351 IFAECI), Buenos Aires, Argentina
extra_authors: 
  - name: Carolina Vera
    affiliation: |
     Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos, Buenos Aires, Argentina
     CONICET – Universidad de Buenos Aires, Centro de Investigaciones del Mar y la Atmósfera (CIMA), Buenos Aires, Argentina
     CNRS – IRD – CONICET – UBA, Instituto Franco‐Argentino para el Estudio del Clima y sus Impactos (UMI 3351 IFAECI), Buenos Aires, Argentina
abstract: |
  Enter the text of your abstract here. This is a sample American Meteorological Society (AMS) \LaTeX\ template.  This document provides authors with instructions on the use of the AMS \LaTeX\ template.  Authors should refer to the file amspaper.tex to review the actual \LaTeX\ code used to create this document. The template.tex file should be modified by authors for their own manuscript.
statement: |
  This is significant becasue I wrote it.
bibliography: ["AsymSAM.bib","packages.bib"]
output: 
  bookdown::pdf_book: 
    base_format: rticles::ams_article
    latex_engine: "xelatex"
  bookdown::word_document2: default
header-includes: 
 - \usepackage{gensymb}
 - \usepackage{subfig}
 - \usepackage[inline]{showlabels}
 - \usepackage[hidelinks]{hyperref}
---

### 2020-10-10 analysis/paper/paper.tex
render

\usepackage{gensymb}
\usepackage{subfig}
\usepackage[inline]{showlabels}

@@

Even though these indices are based on zonal averages, spatial anomalies of geopotential associated with these indices (either by regressino or composition) contain noticeable deviations from zonal symmetry, particularly in the Pacific Ocean region. These zonal asymmetries have not been widely studied, but previous work suggest that they strongly modulate the regional impacts of the SAM \citep[\citet{fogt2012}, \citet{rosso2018}]{fan2007}, going as far as reversing its relationship between precipitation in South America \citep{silvestri2009}. At the very least, the fact that the SAM is not entirely zonally symmetric hinders our ability to reconstruct its historical variability prior to the availability of dense observations in the Southern Hemisphere \citep{jones2009}.

@@

Our objective is, then, to systematically characterise the zonally asymmetric and symmetric components of the SAM variability. For each level, we construct two indices which aim to capture exclusively the variability of the symmetric and asymmetric component respectively. We assess their vertical structure and coherence, temporal variability and trends. We then study the spatial patterns described by the variability exclusive to each index. Finally, we investigate the relationship of the SAM at 700 hPa with temperature and precipitation anomalies.

\#LEAN: Te había hecho el comentario antes pero quizás no había sido del todo claro. Me parece que acá hay que distinguir más en cómo se trabaja metodológicamente cada parte con respecto a los niveles atmosféricos. Por ej. para la evaluación de los impactos solamente se está considerando el ínidice troposférico.

In the Section \ref{sec:methods} we describe the methods. In Section \ref{sec:results}.\ref{sec:temporal} we describe the temporal variability and vertical coherence of the indices. In Section \ref{sec:results}.\ref{sec:spatial}, we analyse the spatial patterns of geopotential height associated with them. In Section \ref{sec:results}.\ref{sec:impacts}, we study their relationship with surface-level temperature and precipitation.

\section{Methods}

\label{sec:methods}

@@

\section{Results}

\label{sec:results}

\subsection{Temporal evolution}

\label{sec:temporal}

@@

\subsection{Spatial patterns}

\label{sec:spatial}

@@

\subsection{Surface Impacts}

\label{sec:impacts}

>>>>

\usepackage{gensymb}
\usepackage{subfig}
\usepackage[inline]{showlabels}
\usepackage[hidelinks]{hyperref}

@@

Even though these indices are based on zonal averages, spatial anomalies of geopotential associated with them (either by regression or composition) contain noticeable deviations from zonal symmetry, particularly in the Pacific Ocean region. These zonal asymmetries have not been widely studied, but previous work suggest that they strongly modulate the regional impacts of the SAM \citep{fan2007, fogt2012, rosso2018}, going as far as reversing its relationship between precipitation in South America \citep{silvestri2009}. At the very least, the fact that the SAM is not entirely zonally symmetric hinders our ability to reconstruct its historical variability prior to the availability of dense observations in the Southern Hemisphere \citep{jones2009}.

@@

Our objective is, then, to systematically characterise the zonally asymmetric and symmetric components of the SAM variability. For each level, we construct two indices which aim to capture exclusively the variability of the symmetric and asymmetric component respectively. We assess their vertical structure and coherence, temporal variability and trends. We then study the spatial patterns described by the variability exclusive to each index focusing on 50 hPa as representing the stratosphere and 700 hPa as representing the troposphere. Finally, we investigate the relationship of the SAM at 700 hPa with temperature and precipitation anomalies.

In the Section \ref{methods} we describe the methods. In Section \ref{temporal} we describe the temporal variability and vertical coherence of the indices. In Section \ref{spatial}, we analyse the spatial patterns of geopotential height associated with them. In Section \ref{impacts}, we study their relationship with surface-level temperature and precipitation.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{temporal}{%
\subsection{Temporal evolution}\label{temporal}}

@@

\hypertarget{spatial}{%
\subsection{Spatial patterns}\label{spatial}}

@@

\hypertarget{impacts}{%
\subsection{Surface Impacts}\label{impacts}}
